Structured Synchronous Reactive Programming for Game Development \\ \Large{Case Study: On Rewriting Pingus from C++ to \CEU}

### 2018-07-12 paper.tex
(>) 6->4 patterns

\begin{abstract}
We present a qualitative case study of rewriting the video game Pingus from C++
to the structured synchronous reactive language \CEU.
%
\CEU supports reactive control-flow primitives that eliminate callbacks and let
programmers write code in direct and sequential style.
Structured reactivity helps describing complex control-flow relationships in
the game logic more concisely.
%
We show gains in productivity for six behaviors in Pingus through a qualitative
analysis of the proposed implementations in \CEU in comparison to the originals
in C++.
%
We also categorize the behaviors in four recurrent control-flow patterns that
likely apply to most games.
\end{abstract}

@@

Our case study shows gains in productivity for six selected behaviors in the
game logic of Pingus rewritten in \CEU.
We present an in-depth qualitative analysis of the proposed solutions in
comparison to the original implementations in C++.
%
Not all techniques result in reduction of \locs (especially considering the
verbose syntax of \CEU), but have other effects such as eliminating shared
variables and dependencies between classes.
%
We also identify four control-flow patterns that likely apply to most games:
        \emph{Finite State Machines},
        \emph{Continuation Passing},
        \emph{Dispatching Hierarchies}, and
        \emph{Lifespan Hierarchies}.
%
A control-flow pattern is a recurring technique to describe execution
dependency and/or explicit ordering between statements.
% (or groups of statements).
%For instance, consider how a key press stimulus propagates through the game
%entities and also what happens with them if the stimulus causes the end of the
%game.
%We focus on a qualitative analysis for the programming techniques that we
%applied during the rewriting process.

@@

The rest of the paper is organized as follows:
Section~\ref{sec.codebase} gives an overview of the Pingus codebases in C++ and
\CEU and describes our approach to identify and rewrite the control flow in the
game.
Section~\ref{sec.pats} discusses six case studies in detail which are
categorized in four control-flow patterns.
Section~\ref{sec.related} discusses related work.
Section~\ref{sec.conclusion} concludes the paper.

@@

We describe six representative game behaviors in detail distributed in the four
patterns and analyze their implementations in C++ and \CEU.%
%\footnote{Due to space constraints, we omit three other game behaviors and a
%fifth pattern \emph{Signaling Mechanisms}.}

@@

We advocate \emph{Structured Synchronous Reactive Programming} as a productive
alternative for game logic development.
%
We use the video game \emph{Pingus} as a qualitative case study.
We compare the implementation of six game behaviors in C++ and \CEU and discuss
how structured reactive mechanisms can eliminate callbacks and let programmers
write code in direct style.
%
Ultimately, we rewrote about 1/4 of the whole codebase (9.186 from 39.362 lines
of code) which comprises the core of the game logic that is susceptible to
structured reactive programming.
%We employed a \emph{live code rewrite}, i.e., starting from the original
%codebase in C++, we reimplemented it piece-by-piece in \CEU without breaking
%the game compilation and execution.

>>>>

\begin{abstract}
We present a qualitative case study of rewriting the video game Pingus from C++
to the structured synchronous reactive language \CEU.
%
\CEU supports reactive control-flow primitives that eliminate callbacks and let
programmers write code in direct and sequential style.
Structured reactivity helps describing complex control-flow relationships in
the game logic more concisely.
%
We show gains in productivity for four behaviors in Pingus through a
qualitative analysis of the proposed implementations in \CEU in comparison to
the originals in C++.
%
We also categorize the behaviors in recurrent control-flow patterns that likely
apply to most games.
\end{abstract}

@@

Our case study shows gains in productivity for four selected behaviors in the
game logic of Pingus rewritten in \CEU.
We present an in-depth qualitative analysis of the proposed solutions in
comparison to the original implementations in C++.
%
Not all techniques result in reduction of \locs (especially considering the
verbose syntax of \CEU), but have other effects such as eliminating shared
variables and dependencies between classes.
%
We also identify four control-flow patterns that likely apply to most games:
        \emph{Finite State Machines},
        \emph{Continuation Passing},
        \emph{Dispatching Hierarchies}, and
        \emph{Lifespan Hierarchies}.
%
A control-flow pattern is a recurring technique to describe execution
dependency and/or explicit ordering between statements.
% (or groups of statements).
%For instance, consider how a key press stimulus propagates through the game
%entities and also what happens with them if the stimulus causes the end of the
%game.
%We focus on a qualitative analysis for the programming techniques that we
%applied during the rewriting process.

@@

The rest of the paper is organized as follows:
Section~\ref{sec.codebase} gives an overview of the Pingus codebases in C++ and
\CEU and describes our approach to identify and rewrite the control flow in the
game.
Section~\ref{sec.pats} discusses four case studies in detail which are
categorized in recurrent control-flow patterns.
Section~\ref{sec.related} discusses related work.
Section~\ref{sec.conclusion} concludes the paper.

@@

We describe representative game behaviors in detail distributed in the four
patterns and analyze their implementations in C++ and \CEU.%
%\footnote{Due to space constraints, we omit three other game behaviors and a
%fifth pattern \emph{Signaling Mechanisms}.}

@@

We advocate \emph{Structured Synchronous Reactive Programming} as a productive
alternative for game logic development.
%
We use the video game \emph{Pingus} as a qualitative case study.
We compare the implementation of four game behaviors in C++ and \CEU and
discuss how structured reactive mechanisms can eliminate callbacks and let
programmers write code in direct style.
%
Ultimately, we rewrote about 1/4 of the whole codebase (9.186 from 39.362 lines
of code) which comprises the core of the game logic that is susceptible to
structured reactive programming.
%We employed a \emph{live code rewrite}, i.e., starting from the original
%codebase in C++, we reimplemented it piece-by-piece in \CEU without breaking
%the game compilation and execution.
